ODGI: understanding pangenome graphs

### 2021-10-06 manuscript/main.tex
towards a nice metrics figure
metrics figure and section ready

% viz / draw
These provide scalable ways of geneating human-viewable pictures of the high-level structure of the graph.
% interesting that we don't go base-level at all... we use vg for that

@@

Applying \texttt{odgi stats}, users can retrieve metrics describing the graph properties, such as the number of nodes, edges, paths, and graph length. MultiQC's \citep{Ewels_2016} ODGI module\footnote{\url{https://multiqc.info/docs/\#odgi} (accessed Oct 2021)} provides an interactive way to comparatively explore the statistics of an arbitrary number of graphs. ODGI also offers more advanced tools for the interrogation of the graphs. \texttt{odgi bin} summarizes the path information into bins of a specified size, enabling a summarized view of gigabase scale graphs. This concept found its way into the pangenome graph ontology \cite{Yokoyama2020}. \texttt{odgi depth} returns the node depth as defined by query criteria, allowing users to assess the complexity of regions in the graph. Collapsed sequences with highly identical repeats indicate intricate loci. Complex motifs can be also detected with \texttt{odgi degree}, which returns the node degree as defined by query criteria. High degree nodes are the mirror of misassemblies or problems in the pangenome building, making the tool useful for validation. In Fig. \ref{fig:metrics}, the characteristics of a human pangenome graph of the exon 1 huntingtin gene \citep{Sathasivam2013,Neueder2017} (\textit{HTTexon1}), obtained with the ODGI metrics tools, are shown.

>>>>

% viz / draw
These provide scalable ways of generating human-viewable pictures of the high-level structure of the graph.
% interesting that we don't go base-level at all... we use vg for that, or STM

@@

Applying \texttt{odgi stats}, users can retrieve metrics describing the graph properties, such as the number of nodes, edges, paths, and graph length. MultiQC's \citep{Ewels_2016} ODGI module\footnote{\url{https://multiqc.info/docs/\#odgi} (accessed Oct 2021)} provides an interactive way to comparatively explore the statistics of an arbitrary number of graphs. ODGI also offers more advanced tools for the interrogation of the graphs. \texttt{odgi bin} summarizes the path information into bins of a specified size, enabling a summarized view of gigabase scale graphs. This concept found its way into the pangenome graph ontology \cite{Yokoyama2020}. \texttt{odgi depth} returns the node depth as defined by query criteria, allowing users to assess the complexity of regions in the graph. Collapsed sequences with highly identical repeats indicate intricate loci. Complex motifs can be also detected with \texttt{odgi degree}, which returns the node degree as defined by query criteria. High degree nodes can be the mirror of variation, misassemblies, or problems in the pangenome building, making the tool useful for validation. In Fig. \ref{fig:metrics}, the characteristics of a human pangenome graph of the exon 1 huntingtin gene \citep{Sathasivam2013,Neueder2017} (\textit{HTTexon1}), obtained with the ODGI metrics tools, are shown.
